Commutative character sheaves and geometric types for supercuspidal representations

### 2017-08-21 Geometric types.tex
Fixing 4.1

If $G$ is nonabelian or disconnected, a similar strategy sometimes succeeds in geometrizing characters $\chi : G(k) \to \EEx$.
Suppose that
\begin{enumerate}
\labitem{(L.1)}{L.1} $G(k) \subset \image(L)$ (which will hold if $G$ is connected).
\end{enumerate}
Then we may associate to each $\chi$ a function $\tilde{\chi}$ as follows.
Define $\tilde{L} : G \to G$ by $\tilde{L}(y) = \Frob{G}(y)y^{-1}$.
If $x \in G(k)$, write $x = \tilde{L}(y)$ and define $\tilde{\chi}(x) = \chi(L(y))$.
Note that $\tilde{\chi}$ is well defined since changing $y$ to $ya$ for $a \in G(k)$ has the
effect of conjugating $L(y)$ by $a^{-1}$, which has no effect on the value $\chi(L(y))$.
In order to obtain a local system with trace $\chi$, we must also assume that
\begin{enumerate}
\labitem{(L.2)}{L.2} $\tilde{\chi}$ is also a character.
\end{enumerate}

\begin{proposition}
Let $G$ be a smooth group scheme over $k$ and assume \ref{L.1} and \ref{L.2}.  Let $\cs{L} = (L_!(\EE)_G)_{\tilde{\chi}^{-1}}$.  Then $\TrFrob{}(\cs{L}) = \chi$.
\end{proposition}
\begin{proof}
The stalk of $\cs{L}$ at $x \in G(k)$ is
\[
\cs{L}_x = \{ s : L^{-1}(x) \to \EE : s(ay) = \tilde{\chi}^{-1}(a) s(y) \mbox{ for all $a \in G(k), y \in L^{-1}(x)$}\}.
\]
Frobenius acts by
\[
(^{\Frob{}}s)(y) = s(\Frob{}^{-1}(y)) = s(yx^{-1}) = s(yx^{-1}y^{-1}y) = \tilde{\chi}(yxy^{-1})s(y) = \tilde{\chi}(\tilde{L}(y))s(y) = \chi(x)s(y),
\]
thus proving the claim. Need $z$ with $L(z) = \tilde{L}(y)$.  Note that $z = \Frob{G}(y^{-1})$ does not do the job.
\end{proof}

>>>>

If $G$ is nonabelian or disconnected, a similar strategy sometimes succeeds in geometrizing characters $\chi : G(k) \to \EEx$.
Define $\tilde{L} : G \to G$ by $\tilde{L}(y) = \Frob{G}(y)y^{-1}$.
The fibers of $L$ are right cosets of $G(k)$ and the fibers of $\tilde{L}$ are left cosets.
Suppose that
\begin{enumerate}
\labitem{(L.1)}{L.1} $G(k) \subset \image(\tilde{L}) \cap \image(L)$ (which will hold if $G$ is connected).
\end{enumerate}
Then we may associate to each $\chi$ a function $\tilde{\chi} : G(k) \to \EEx$ as folllows.
If $z \in G(k)$, write $z = \tilde{L}(y)$ and define $\tilde{\chi}(z) = \chi(L(y))$.
Note that $\tilde{\chi}$ is well defined since changing $y$ to $ya$ for $a \in G(k)$ has the
effect of conjugating $L(y)$ by $a^{-1}$, which has no effect on the value $\chi(L(y))$.
In order to obtain a local system with trace $\chi$, we must also assume that
\begin{enumerate}
\labitem{(L.2)}{L.2} $\tilde{\chi}$ is also a character.
\end{enumerate}

\begin{proposition}
Let $G$ be a smooth group scheme over $k$ and assume \ref{L.1} and \ref{L.2}.  Let $\cs{L} = (L_!(\EE)_G)_{\tilde{\chi}^{-1}}$.  Then $\TrFrob{}(\cs{L}) = \chi$.
\end{proposition}
\begin{proof}
The stalk of $\cs{L}$ at $x \in G(k)$ is
\[
\cs{L}_x = \{ s : L^{-1}(x) \to \EE : s(ay) = \tilde{\chi}^{-1}(a) s(y) \mbox{ for all $a \in G(k), y \in L^{-1}(x)$}\}.
\]
For any $y \in L^{-1}(x)$, note that $\Frob{G}^{-1}(y) = yx^{-1}$ and $\tilde{L}(y) = yxy^{-1} \in G(k)$.  Frobenius acts on $s \in \cs{L}_x$ by
\[
(^{\Frob{}}s)(y) = s(\Frob{G}^{-1}(y)) = \tilde{\chi}^{-1}(yx^{-1}y^{-1})s(y) = \tilde{\chi}(\tilde{L}(y))s(y) = \chi(x)s(y),
\]
thus proving the claim.
\end{proof}
